### 2014-03-11 thesis/src/chapters/chapter-state-of-the-art.tex
Thesis: degradation of ch3

\chapter{Related work}

@@

\section{Requirements}
One can notice that elements that yield a solution to a problem stated in the first chapter, which is ensuring that users' application provide appropriate Quality-of-Service for its customers in a most-cost effective manner, were gradually introduced in the previous chapter:

@@

\subsection{Functional requirements}
\begin{enumerate}
  \item The user of the platform is able to:
    \begin{enumerate}
      \item deploy a service,
      \item cancel the service,
      \item check the status of the previously ordered-to-deploy service at any time. \emph{Status} means 
        \begin{inparaenum}[a)]
        \item whether or not the deployment succeeded,
        \item current uptime of the service,
        \item current cost
        \end{inparaenum}
    \end{enumerate}
  \item One of the elements of the platform is a client application that is used by the user of the platform to communicate with it
  \item During the deployment, the platform takes as an input a description of the service (application) that consists of:
    \begin{itemize}
    \item service name,
    \item software stacks (e.g. \emph{java}, \emph{ruby}),
    \item scaling policies (per each stack) which define
      \begin{inparaenum}[i)]
      \item minimal and maximal number of VMs that are needed for the stack,
      \item name of the policy (algorithm) used for scaling,
      \item parameters of the policy
      \end{inparaenum}
    \end{itemize}
  \item Deployment of a service is done in a way which minimises the cost from the client's perspective with ensuring Quality-of-Service requirements at the same time,
  \item The platform monitors the state of the deployed services and based on the results of this process takes appropriate steps in order to meet the auto-scaling requirements. These include
    \begin{inparaenum}[a)]
    \item altering VM's parameters and configuration,
    \item vertical scaling,
    \item horizontal scaling,
    \item scaling stacks among different cloud providers
    \end{inparaenum}
\end{enumerate}

\subsection{Non-functional}
\begin{itemize}
  \item The platform does not confine itself to one provider, but to a \emph{ecosystem of various cloud providers} that offers deployment capabilities which vary in terms of quality of service, cost, etc.
\end{itemize}

% radek
\section{Carina}

\subsection{Introduction}
Carina is an open source project, released under Apache License 2.0 and built on top of OpenNebula. It aims to ``(\ldots) standardize the process for automating multi-VM deployments and setting auto-scaling and availability management policies in the cloud.'' \cite{CarinaBlog}. The project is used by the authors at their work at RIM in an OpenNebula-based private cloud.

\subsection{Features}
As it is stated in the requirements of the solution, \emph{Carina} should support variety of features which can be considered worth scrutinizing carefully as they are closely related to notions of adaptivity and scalability. To name the most relevant: \cite{CarinaBlog}
\begin{itemize}
  \item Collect and aggregate OS or app-specific metrics across a cluster
  \item Drive elastic scaling of clusters based on workload or events
  \item Support deployment and handling of failover of services across multiple datacenters
\end{itemize}

@@

\subsection{Key technical concepts}

\subsubsection{Adaptivity}
In \emph{Carina} there are mechanisms which can perform application scaling, both in manual and automatic fashion. It is possible for a system administrator to directly modify the existing application and its environment by changing its parameters manually. Automatic management of an environment is done by defining \emph{scaling policies} which can be in two flavours: \emph{time-based} and \emph{load-based}. Those are called by the authors \emph{elasticy policies}.

@@

\subsubsection{Scaling}
Section 2.2 introduced and discussed various types of application scaling. \emph{Carina} offers only one type of scaling and this is \emph{horizontal} scaling -- user's capabilities are limited only to add/remove a virtual machine to/from a given environment. Unfortunately there are no means to modify the parameters of a given or a set of virtual machines.

\subsubsection{Cloud federation awareness}
In this paper we strongly advocate spanning cloud resources across multiple vendors/providers. Despite having in its requirements such a position, \emph{Carina} \textbf{does not} support deployment or scaling of services across different providers. In the environment configuration we define only \textbf{one} endpoint -- it can be thought as a reference to a cloud provider. As it can be only one per environment, it is not possible to deploy or scale artifacts across many providers.

\subsection{Design}
To complete the discussion about \emph{Carina} it is essential to include its architecture overview. This is shown in figure \ref{sota:carina-design}.

@@

\subsection{Summary}
\emph{Carina} emerged as a promising platform built on top of OpenNebula, which lacked most of its functionalities at that time. The good point is the fact that this product is used in a business context and judging by the authors' description it actually works. However, due to deficiencies mentioned in the previous sections, \emph{Carina} cannot be considered a mature and generic solution for more demanding cloud providers.

\section{OneFlow}

\subsection{Introduction}
\emph{OneFlow} \cite{OneFlow} is a management system integrated into \emph{OpenNebula} that aspires to enhance it with multi-tiered applications (service) deployment and auto-scaling capabilities. It is available as a core part of \emph{OpenNebula} since 4.2 version and released under \emph{Apache License, Version 2.0}.

\subsection{Features}
In a context of this thesis, \emph{OneFlow} enriches \emph{OpenNebula} with only one, yet powerful feature: application scaling based on policies. Besides this, benefits of using \emph{OneFlow} are as follows:
\begin{itemize}
\item defining multi-tiered applications (services) as a collection of applications
\item providing configurable services from a catalogue and self-service portal
\item enabling tight, efficient administrative control
\item fine-grained access control for the secure sharing of services with other users
\end{itemize}

\subsection{Key technical concepts}
\subsubsection{Adaptivity}
Adaptivity is achieved by specifying policies, that violated, trigger events impacting service environment. There are two types of policies available:
\begin{itemize}
 \item \emph{Auto-scaling based on metrics} - policy defines an expression that triggers scaling adjustments. Expressions can use performance data sent from virtual machine using \emph{OneGate} as well as virtual machine data such as CPU, memory, network bandwidth
 \item \emph{Auto-scaling based on schedule} - policy specifies a time or a time recurrence and correlates that information to appropriate service adjustments
\end{itemize}

@@

\subsubsection{Scaling}
\emph{OneFlow} is capable of \textbf{only} horizontal scaling: it adds or removes virtual machine using \emph{OpenNebula API}. Violating given policy results in virtual machine's state transition into \texttt{SCALING}. After proper action is finished, virtual machines converges to a \texttt{COOLDOWN} state. What this state transition does is to allow an environment to accommodate changes and adjust to a recent changes, disabling scaling for a given period of time.

\subsubsection{Cloud federation awareness}
\emph{OneFlow} itself \textbf{does not} cooperate with different cloud instance. Hence, one is limited to \emph{OpenNebula} hybrid or public cloud capabilities that employs oZones and \emph{AWS}, respectively. Although it expands single \emph{OpenNebula} instance capacity, it is limited in a way that discriminates it as possible solution. Firstly, the problem with oZones is that they require centralised management, while cloud federation is by definition distributed and independent. Secondly, cloud bursting into \emph{AWS} limits instance monitoring capabilities. Consequently, it is not possible to supervise provisioned instances.

@@

\section{OpenShift}

\subsection{Introduction}
\emph{OpenShift} is a developed by \emph{RedHat} Platform-as-a-Service solution. It is available in two different flavours:
\begin{itemize}
 \item \emph{OpenShift Enterprise} (private cloud)
 \item \emph{OpenShift Online / FreeShift} hosted by RedHat (public cloud)
\end{itemize}

@@

\subsection{Features}
High level features that distinguishes \emph{OpenShift} are as follows:
\begin{itemize}
  \item accelerated application service delivery by on-demand and self-provision application stack access
  \item minimised vendor lock-in by portability (there are no proprietary APIs, technologies)
  \item automatic application scaling
\end{itemize}

@@

\subsection{Key technical concepts}

\subsubsection{Adaptivity}
\emph{OpenShift}'s adaptivity capabilities are constrained to a simple case: while creating a new application, it is possible to specify whether application should scale. As a consequence, additional gear is allocated to serve \emph{HAProxy} \cite{HAProxy}. There is only \textbf{one built-in} scaling policy that is based on a relatively straightforward algorithm \cite{OpenShiftScaling}:
\emph{\begin{quote}
The algorithm for scaling up and scaling down is based on the number of concurrent requests to your application. OpenShift allocates 10 connections per gear - if HAProxy sees that you're sustaining 90\% of your peak capacity, it adds another gear. If your demand falls to 50\% of your peak capacity for several minutes, HAProxy removes that gear.
\end{quote}}

\subsubsection{Scalability}
As previous point implies, OpenShift \textbf{solely} uses horizontal scaling: it adds additional gears if number of requests increases beyond a certain level.

\subsubsection{Cloud federation awareness}
OpenShift does not use any notion of a cloud provider or datacenter. However, since cartridges are in fact Linux processes running on OpenShift Nodes, it is possible to deploy  OpenShift Enterprise on any infrastructure running RHEL. It can consist of physical nodes as well as virtual machines, managed, for example, by OpenStack. With all that said, is still requires a centralised, OpenShift controlled environment. Hence, OpenShift \textbf{is not} capable to work in a cloud federation.

\subsubsection{Multi-tenancy}
In OpenShift, each application is represented as a set of gears. For example, gear is a database or application server. Gears are hosted on OpenShift Nodes. There is a many-to-one relationship between gear and OpenShift Nodes. What is leveraged to achieve this multi-tenancy is:
\begin{itemize}
 \item SELinux, isolation between applications running at the same node
 \item control groups (cgroups), fine-grained control over the memory / CPU / IO utilisation / networking on per process basis
 \item kernel namespaces, groups of processes are separated, so that they cannot see resources in other groups
\end{itemize}

\subsection{Summary}
Summing up, key advantage of the OpenShift seems to be it simplicity: creating and deploying applications using OpenShift is as simple as it gets. Beside this, built in support for most popular technologies and simple API makes OpenShift attractive platform. Having that said, oversimplified auto-scaling and being constrained to a single cloud provider may be not sufficient in complicated scenarios.

% radek
\section{CloudFoundry}
\subsection{Overview}
\emph{Cloud Foundry} is a platform-as-a-service solution which can be installed on local or off-premises infrastructure, such as Amazon Web Services, OpenStack or vSphere. It is available in two versions as
\begin{itemize}
  \item an open-source project
  \item a commercial solution, \emph{Pivotal CF}
\end{itemize}
In this section we want to elaborate only on an open-source project.

\subsection{History}
The project was initially developed by VMWare and its first release was in 2011, It was then that VMware decided to make it available to the general public and released it under Apache License 2.0. For the next consecutive two years, there two versions of \emph{Cloud Foundry} -- as a hosted solution owned by VMWare and an open-source project -- were maintained . However, in December 2012 VMware and EMC shared information on the \emph{Pivotal Initiative} \cite{PivotalInitiative} -- a virtual organization of people from both companies with background in \emph{big data} and \emph{cloud application platforms}. With the advent of the this new entity, \emph{Pivotal} \cite{GoPivotal}, it was announced the launch of the new product -- Pivotal One, powered by \emph{Pivotal CF}, the enterprise version of \emph{Cloud Foundry}. Now users can choose between the open-source and commercial solutions.

\subsection{Features}
Below there are listed some of the main functionalities offered by \emph{Cloud Foundry}:
\begin{itemize}
  \item Support for multi-provider ecosystem (\emph{Multi-Cloud})
  \item Reduced the management burden of the life-cycle of an application by a simplified CLI that enables, e.g. faster and instant deployment,
  \item Dynamic routing which enables and enhances horizontal scaling and updating applications,
  \item Usage of own-developed lightweight containers (\emph{Warden}), which ensures the proper isolation level and improves the speed of movement of application instances between virtual machines (utilizing \emph{aufs} -- advanced multi layered unification filesystem)
  \item Health management -- continually monitoring application instances and taking according recovery actions if needed
  \item Standards-based authorization and authentication system which supports various protocols such as LDAP, SAML or OAuth 2.0
\end{itemize}

\subsection{Key technical concepts}

\subsubsection{Cloud Foundry-specific}
\begin{asparaenum}
\item[\textbf{Warden}] The main goal of \emph{Warden} is to manage isolated and resource-controller environments called containers. It is responsible for management of the whole lifecycle of an environment and provides API for actually performing any operation on a container. At first, \emph{Warden} used LXC technology under the hood, but now containers are an own product of \emph{Cloud Foundry} community.
\item[\textbf{Droplet Execution Agent}] \emph{DEA} manages \emph{Warden} containers, stages user's application and runs droplets -- wrappers around already deployed applications.
\item[\textbf{BOSH}] \emph{BOSH} is a tool chain for release, deployment and lifecycle management of distributed services. Its functionalities can be compared to those offered by \emph{Chef}.
\end{asparaenum}

\subsubsection{Adaptivity}
\emph{Cloud foundry} does not support auto-scaling -- it is only possible for the cloud consumer to scale the application manually. However, there can be additional tools which can add this functionality, such as \emph{RightScale}.

@@

\subsubsection{Scalability}
\emph{Cloud foundry} supports horizontal scaling not only within one provider, but also across many. This feature is called \emph{Multi-Cloud} in Cloud Foundry nomenclature.

\subsubsection{Cloud-federation awareness}
As it was stated in previous sections, Cloud Foundry is \emph{cloud-federation aware}, i.e. it allows users to span their applications across many providers, which results in a hybrid solution that comes in many flavours: \emph{public-public}, \emph{public-private} or \emph{private-private}.

\subsection{Summary}
Owing to the variety of useful features, great support from community and significant IT enterprises such as IBM, VMware to name a few, \emph{Cloud Foundry} can be considered a mature product which can compete with current leaders in the fields, such as Amazon AWS.
Additionally, it is highly probable that acquiring the hosted solution by \emph{Pivotal} may be of a great benefit to the whole cloud community as the market will be more competitive.

% radek
\section{Providers comparison}
In this section we would like to compare all depicted providers in terms of fulfilling functional requirements specified in the overview. Table \ref{tab:cloud-providers-comparision-features} presents such comparison. Table \ref{tab:cloud-providers-adaptivity} takes more detailed approach and summarises how platform tackles adaptation: their data analysis mechanisms and policies support.

>>>>

\section{Related work}

@@

\subsection{Requirements}
One can notice that elements that yield a solution to a problem stated in the first chapter, which is ensuring that users' application provide appropriate Quality-of-Service for its customers in a most-cost effective manner, were gradually introduced in the previous chapter:

@@

\subsubsection{Functional requirements}
\begin{enumerate}
  \item The user of the platform is able to:
    \begin{enumerate}
      \item deploy a service,
      \item cancel the service,
      \item check the status of the previously ordered-to-deploy service at any time. \emph{Status} means 
        \begin{inparaenum}[a)]
        \item whether or not the deployment succeeded,
        \item current uptime of the service,
        \item current cost
        \end{inparaenum}
    \end{enumerate}
  \item One of the elements of the platform is a client application that is used by the user of the platform to communicate with it
  \item During the deployment, the platform takes as an input a description of the service (application) that consists of:
    \begin{itemize}
    \item service name,
    \item software stacks (e.g. \emph{java}, \emph{ruby}),
    \item scaling policies (per each stack) which define
      \begin{inparaenum}[i)]
      \item minimal and maximal number of VMs that are needed for the stack,
      \item name of the policy (algorithm) used for scaling,
      \item parameters of the policy
      \end{inparaenum}
    \end{itemize}
  \item Deployment of a service is done in a way which minimises the cost from the client's perspective with ensuring Quality-of-Service requirements at the same time,
  \item The platform monitors the state of the deployed services and based on the results of this process takes appropriate steps in order to meet the auto-scaling requirements. These include
    \begin{inparaenum}[a)]
    \item altering VM's parameters and configuration,
    \item vertical scaling,
    \item horizontal scaling,
    \item scaling stacks among different cloud providers
    \end{inparaenum}
\end{enumerate}

\subsubsection{Non-functional}
\begin{itemize}
  \item The platform does not confine itself to one provider, but to a \emph{ecosystem of various cloud providers} that offers deployment capabilities which vary in terms of quality of service, cost, etc.
\end{itemize}

% radek
\subsection{Carina}

\subsubsection{Introduction}
Carina is an open source project, released under Apache License 2.0 and built on top of OpenNebula. It aims to ``(\ldots) standardize the process for automating multi-VM deployments and setting auto-scaling and availability management policies in the cloud.'' \cite{CarinaBlog}. The project is used by the authors at their work at RIM in an OpenNebula-based private cloud.

\subsubsection{Features}
As it is stated in the requirements of the solution, \emph{Carina} should support variety of features which can be considered worth scrutinizing carefully as they are closely related to notions of adaptivity and scalability. To name the most relevant: \cite{CarinaBlog}
\begin{itemize}
  \item Collect and aggregate OS or app-specific metrics across a cluster
  \item Drive elastic scaling of clusters based on workload or events
  \item Support deployment and handling of failover of services across multiple datacenters
\end{itemize}

@@

\subsubsection{Key technical concepts}

\begin{asparaenum}
\item[\textbf{Adaptivity}]
In \emph{Carina} there are mechanisms which can perform application scaling, both in manual and automatic fashion. It is possible for a system administrator to directly modify the existing application and its environment by changing its parameters manually. Automatic management of an environment is done by defining \emph{scaling policies} which can be in two flavours: \emph{time-based} and \emph{load-based}. Those are called by the authors \emph{elasticy policies}.

@@

\item[\textbf{Scaling}]
Section 2.2 introduced and discussed various types of application scaling. \emph{Carina} offers only one type of scaling and this is \emph{horizontal} scaling -- user's capabilities are limited only to add/remove a virtual machine to/from a given environment. Unfortunately there are no means to modify the parameters of a given or a set of virtual machines.

\item[\textbf{Cloud federation awareness}]
In this paper we strongly advocate spanning cloud resources across multiple vendors/providers. Despite having in its requirements such a position, \emph{Carina} \textbf{does not} support deployment or scaling of services across different providers. In the environment configuration we define only \textbf{one} endpoint -- it can be thought as a reference to a cloud provider. As it can be only one per environment, it is not possible to deploy or scale artifacts across many providers.

\end{asparaenum}

\subsubsection{Design}
To complete the discussion about \emph{Carina} it is essential to include its architecture overview. This is shown in figure \ref{sota:carina-design}.

@@

\subsubsection{Summary}
\emph{Carina} emerged as a promising platform built on top of OpenNebula, which lacked most of its functionalities at that time. The good point is the fact that this product is used in a business context and judging by the authors' description it actually works. However, due to deficiencies mentioned in the previous sections, \emph{Carina} cannot be considered a mature and generic solution for more demanding cloud providers.

\subsection{OneFlow}

\subsubsection{Introduction}
\emph{OneFlow} \cite{OneFlow} is a management system integrated into \emph{OpenNebula} that aspires to enhance it with multi-tiered applications (service) deployment and auto-scaling capabilities. It is available as a core part of \emph{OpenNebula} since 4.2 version and released under \emph{Apache License, Version 2.0}.

\subsubsection{Features}
In a context of this thesis, \emph{OneFlow} enriches \emph{OpenNebula} with only one, yet powerful feature: application scaling based on policies. Besides this, benefits of using \emph{OneFlow} are as follows:
\begin{itemize}
\item defining multi-tiered applications (services) as a collection of applications
\item providing configurable services from a catalogue and self-service portal
\item enabling tight, efficient administrative control
\item fine-grained access control for the secure sharing of services with other users
\end{itemize}

\subsubsection{Key technical concepts}

\begin{asparaenum}

\item[\textbf{Adaptivity}]
Adaptivity is achieved by specifying policies, that violated, trigger events impacting service environment. There are two types of policies available:
\begin{itemize}
 \item \emph{Auto-scaling based on metrics} - policy defines an expression that triggers scaling adjustments. Expressions can use performance data sent from virtual machine using \emph{OneGate} as well as virtual machine data such as CPU, memory, network bandwidth
 \item \emph{Auto-scaling based on schedule} - policy specifies a time or a time recurrence and correlates that information to appropriate service adjustments
\end{itemize}

@@

\item[\textbf{Scaling}]
\emph{OneFlow} is capable of \textbf{only} horizontal scaling: it adds or removes virtual machine using \emph{OpenNebula API}. Violating given policy results in virtual machine's state transition into \texttt{SCALING}. After proper action is finished, virtual machines converges to a \texttt{COOLDOWN} state. What this state transition does is to allow an environment to accommodate changes and adjust to a recent changes, disabling scaling for a given period of time.

\item[\textbf{Cloud federation awareness}]
\emph{OneFlow} itself \textbf{does not} cooperate with different cloud instance. Hence, one is limited to \emph{OpenNebula} hybrid or public cloud capabilities that employs oZones and \emph{AWS}, respectively. Although it expands single \emph{OpenNebula} instance capacity, it is limited in a way that discriminates it as possible solution. Firstly, the problem with oZones is that they require centralised management, while cloud federation is by definition distributed and independent. Secondly, cloud bursting into \emph{AWS} limits instance monitoring capabilities. Consequently, it is not possible to supervise provisioned instances.
\end{asparaenum}

@@

\subsection{OpenShift}

\subsubsection{Introduction}
\emph{OpenShift} is a developed by \emph{RedHat} Platform-as-a-Service solution. It is available in two different flavours:
\begin{itemize}
 \item \emph{OpenShift Enterprise} (private cloud)
 \item \emph{OpenShift Online / FreeShift} hosted by RedHat (public cloud)
\end{itemize}

@@

\subsubsection{Features}
High level features that distinguishes \emph{OpenShift} are as follows:
\begin{itemize}
  \item accelerated application service delivery by on-demand and self-provision application stack access
  \item minimised vendor lock-in by portability (there are no proprietary APIs, technologies)
  \item automatic application scaling
\end{itemize}

@@

\subsubsection{Key technical concepts}
\begin{asparaenum}

\item[\textbf{Adaptivity}]
\emph{OpenShift}'s adaptivity capabilities are constrained to a simple case: while creating a new application, it is possible to specify whether application should scale. As a consequence, additional gear is allocated to serve \emph{HAProxy} \cite{HAProxy}. There is only \textbf{one built-in} scaling policy that is based on a relatively straightforward algorithm \cite{OpenShiftScaling}:
\emph{\begin{quote}
The algorithm for scaling up and scaling down is based on the number of concurrent requests to your application. OpenShift allocates 10 connections per gear - if HAProxy sees that you're sustaining 90\% of your peak capacity, it adds another gear. If your demand falls to 50\% of your peak capacity for several minutes, HAProxy removes that gear.
\end{quote}}

\item[\textbf{Scalability}]
As previous point implies, OpenShift \textbf{solely} uses horizontal scaling: it adds additional gears if number of requests increases beyond a certain level.

\item[\textbf{Cloud federation awareness}]
OpenShift does not use any notion of a cloud provider or datacenter. However, since cartridges are in fact Linux processes running on OpenShift Nodes, it is possible to deploy  OpenShift Enterprise on any infrastructure running RHEL. It can consist of physical nodes as well as virtual machines, managed, for example, by OpenStack. With all that said, is still requires a centralised, OpenShift controlled environment. Hence, OpenShift \textbf{is not} capable to work in a cloud federation.

\item[\textbf{Multi-tenancy}]
In OpenShift, each application is represented as a set of gears. For example, gear is a database or application server. Gears are hosted on OpenShift Nodes. There is a many-to-one relationship between gear and OpenShift Nodes. What is leveraged to achieve this multi-tenancy is:
\begin{itemize}
 \item SELinux, isolation between applications running at the same node
 \item control groups (cgroups), fine-grained control over the memory / CPU / IO utilisation / networking on per process basis
 \item kernel namespaces, groups of processes are separated, so that they cannot see resources in other groups
\end{itemize}

\end{asparaenum}

\subsubsection{Summary}
Summing up, key advantage of the OpenShift seems to be it simplicity: creating and deploying applications using OpenShift is as simple as it gets. Beside this, built in support for most popular technologies and simple API makes OpenShift attractive platform. Having that said, oversimplified auto-scaling and being constrained to a single cloud provider may be not sufficient in complicated scenarios.

% radek
\subsection{CloudFoundry}
\subsubsection{Overview}
\emph{Cloud Foundry} is a platform-as-a-service solution which can be installed on local or off-premises infrastructure, such as Amazon Web Services, OpenStack or vSphere. It is available in two versions as
\begin{itemize}
  \item an open-source project
  \item a commercial solution, \emph{Pivotal CF}
\end{itemize}
In this section we want to elaborate only on an open-source project.

\subsubsection{History}
The project was initially developed by VMWare and its first release was in 2011, It was then that VMware decided to make it available to the general public and released it under Apache License 2.0. For the next consecutive two years, there two versions of \emph{Cloud Foundry} -- as a hosted solution owned by VMWare and an open-source project -- were maintained . However, in December 2012 VMware and EMC shared information on the \emph{Pivotal Initiative} \cite{PivotalInitiative} -- a virtual organization of people from both companies with background in \emph{big data} and \emph{cloud application platforms}. With the advent of the this new entity, \emph{Pivotal} \cite{GoPivotal}, it was announced the launch of the new product -- Pivotal One, powered by \emph{Pivotal CF}, the enterprise version of \emph{Cloud Foundry}. Now users can choose between the open-source and commercial solutions.

\subsubsection{Features}
Below there are listed some of the main functionalities offered by \emph{Cloud Foundry}:
\begin{itemize}
  \item Support for multi-provider ecosystem (\emph{Multi-Cloud})
  \item Reduced the management burden of the life-cycle of an application by a simplified CLI that enables, e.g. faster and instant deployment,
  \item Dynamic routing which enables and enhances horizontal scaling and updating applications,
  \item Usage of own-developed lightweight containers (\emph{Warden}), which ensures the proper isolation level and improves the speed of movement of application instances between virtual machines (utilizing \emph{aufs} -- advanced multi layered unification filesystem)
  \item Health management -- continually monitoring application instances and taking according recovery actions if needed
  \item Standards-based authorization and authentication system which supports various protocols such as LDAP, SAML or OAuth 2.0
\end{itemize}

\subsubsection{Key technical concepts}

\begin{asparaenum}

\item[\textbf{Cloud Foundry-specific}] Technical solutions that makes CF interesting are as follows:
\begin{itemize}
\item Warden - The main goal of \emph{Warden} is to manage isolated and resource-controller environments called containers. It is responsible for management of the whole lifecycle of an environment and provides API for actually performing any operation on a container. At first, \emph{Warden} used LXC technology under the hood, but now containers are an own product of \emph{Cloud Foundry} community.
\item Droplet Execution Agent - \emph{DEA} manages \emph{Warden} containers, stages user's application and runs droplets -- wrappers around already deployed applications.
\item BOSH - \emph{BOSH} is a tool chain for release, deployment and lifecycle management of distributed services. Its functionalities can be compared to those offered by \emph{Chef}.
\end{itemize}

\item[\textbf{Adaptivity}]
\emph{Cloud foundry} does not support auto-scaling -- it is only possible for the cloud consumer to scale the application manually. However, there can be additional tools which can add this functionality, such as \emph{RightScale}.

@@

\item[\textbf{Scalability}]
\emph{Cloud foundry} supports horizontal scaling not only within one provider, but also across many. This feature is called \emph{Multi-Cloud} in Cloud Foundry nomenclature.

\item[\textbf{Cloud-federation awareness}]
As it was stated in previous sections, Cloud Foundry is \emph{cloud-federation aware}, i.e. it allows users to span their applications across many providers, which results in a hybrid solution that comes in many flavours: \emph{public-public}, \emph{public-private} or \emph{private-private}.
\end{asparaenum}

\subsubsection{Summary}
Owing to the variety of useful features, great support from community and significant IT enterprises such as IBM, VMware to name a few, \emph{Cloud Foundry} can be considered a mature product which can compete with current leaders in the fields, such as Amazon AWS.
Additionally, it is highly probable that acquiring the hosted solution by \emph{Pivotal} may be of a great benefit to the whole cloud community as the market will be more competitive.

% radek
\subsection{Providers comparison}
In this section we would like to compare all depicted providers in terms of fulfilling functional requirements specified in the overview. Table \ref{tab:cloud-providers-comparision-features} presents such comparison. Table \ref{tab:cloud-providers-adaptivity} takes more detailed approach and summarises how platform tackles adaptation: their data analysis mechanisms and policies support.
